Measurement and Probability in Fuzzy Quantum Theories

### 2017-01-24 prop.tex
prop.tex \section{Increasing the Measurement Resource} Updating...

\noindent It is easy to establish that quantum interval-valued
probability measures generalize conventional quantum probability
measures. In particular, any quantum probability measure can be recast
as an interval-valued measure using the three intervals
$\left[0,0\right]$, $\left[1,1\right]$ and \emph{$\left[0,1\right]$},
where $\left[0,0\right]$ and $\left[1,1\right]$ are called \imposs~and
\necess~as before, and \emph{$\left[0,1\right]$} is called
\unknown~because it provides no information. Given a quantum
probability measure~$\mu:\events\rightarrow\left[0,1\right]$, we
define a quantum interval-valued probability
measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$ by
$\bar{\mu}(P)=\iota\left(\mu(P)\right)$, where
$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ is defined by
\[
\iota(x)=\begin{cases}
\necess & \textrm{if }x=1\textrm{ ;}\\
\imposs & \textrm{if }x=0\textrm{ ;}\\
\unknown & \textrm{otherwise.}
\end{cases}
\]
This measure represents the beliefs of an experimenter with no prior 
knowledge about the particular quantum system in question. Formally,
we can ask: what can we deduce about the state of a quantum system
given a quantum interval-valued probability measure, i.e., given
observations done with finite resources. In the case the intervals are
infinitely precise the question reduces to Gleason's theorem which
states that the state of the quantum system is uniquely determined by
the probability measure. But surely the less resources are available,
the less precise the intervals, and the less we expect to know about
the state of the system. To formally state and answer this question we
begin with defining the \emph{core} and \emph{convexity} of a
probability measure as follows.

@@

To prove the other direction, consider the linear transformation~$T:\mathbb{C}^{3}\rightarrow\mathbb{C}^{2}$
defined by $T\left(\ket{\varphi}\right)=\begin{pmatrix}\ip{\psi}{\varphi}\\
\ip{\phi}{\varphi}
\end{pmatrix}$. Since $\ket{\psi}\nparallel\ket{\phi}$, we have $\mathrm{rank}\left(T\right)=2$.
Together with the rank equation~\cite{FraleighBeauregard1995}, we
have
\[
\dim\left(\ker T\right)=\mathrm{nullity}\left(T\right)=3-\mathrm{rank}\left(T\right)=3-2=1\textrm{ .}
\]
By the other direction of the proof, we have already known $\ket{\psi}\times\ket{\phi}\in\ker T$.
Therefore, $\ket{\varphi}\parallel\left(\ket{\psi}\times\ket{\phi}\right)$.
\end{proof}
Since interval-valued probability measures~$\bar{\mu}$ and interval-valued
frame functions~$\bar{f}$ are one-to-one correspondence, studying
1-dimensional projectors mapping to impossible by $\bar{\mu}$ is
the same as studying the points on $\mathbb{C}S^{2}$ mapping to impossible
by $\bar{f}$. These points, denoted by $\bar{f}^{-1}\left(\imposs\right)=\set{\ket{\psi}\in\mathbb{C}^{3}}{\bar{f}(\ket{\psi})=\imposs}$,
play an important role to understand $\bar{f}$. There are 4 different
cases when analyzing $\bar{f}^{-1}\left(\imposs\right)$:
\begin{enumerate}
\item $\bar{f}^{-1}\left(\imposs\right)=\emptyset$. If $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, we could say nothing about $\ket{\phi}$.
\item $\bar{f}^{-1}\left(\imposs\right)=\left\{ \ket{\psi_{0}}\right\} $.
If $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$ for some pure
state~$\ket{\phi}$, $\ket{\phi}$ could be any state orthogonal
to $\ket{\psi_{0}}$.
\item Consider $\left\{ \ket{\psi_{0}},\ket{\psi_{1}}\right\} \subseteq\bar{f}^{-1}\left(\imposs\right)$.
If $\left(\ket{\psi_{0}}\times\ket{\psi_{1}}\right)\parallel\left(\ket{\psi_{0}'}\times\ket{\psi_{1}'}\right)$
for any $\left\{ \ket{\psi_{0}'},\ket{\psi_{1}'}\right\} \subseteq\bar{f}^{-1}\left(\imposs\right)$,
then $\ket{\psi_{0}}\times\ket{\psi_{1}}$ is the only pure state
which might induce $\bar{f}$. In another word, if $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, then $\ket{\phi}\parallel\left(\ket{\psi_{0}}\times\ket{\psi_{1}}\right)$. 
\item If another pair of states~$\ket{\psi_{0}'}$ and $\ket{\psi_{1}'}\in\bar{f}^{-1}\left(\imposs\right)$
gives different state of cross product~$\ket{\psi_{0}'}\times\ket{\psi_{1}'}$,
then both $\ket{\psi_{0}}\times\ket{\psi_{1}}$ and $\ket{\psi_{0}'}\times\ket{\psi_{1}'}$
might induce $\bar{f}$, but neither $f_{\ket{\psi_{0}}\times\ket{\psi_{1}}}^{B}$
nor $f_{\ket{\psi_{0}'}\times\ket{\psi_{1}'}}^{B}$ in the core of
$\bar{f}$ because $f_{\ket{\psi_{0}}\times\ket{\psi_{1}}}^{B}$ is
inconsistent with $\bar{f}(\proj{\psi_{0}'})=\bar{f}(\proj{\psi_{1}'})=\imposs$
and $f_{\ket{\psi_{0}'}\times\ket{\psi_{1}'}}^{B}$ is inconsistent
with $\bar{f}(\proj{\psi_{0}})=\bar{f}(\proj{\psi_{1}})=\imposs$.
Actually, $\bar{f}$ has an empty core. 
\end{enumerate}
In summary, if we have an idea of $\bar{f}^{-1}\left(\imposs\right)$,
we could have some idea of the pure states in core.

As we did in Fig.~\ref{fig:three-dimensional-4-value}, we could
plot the 1-dimensional projectors in a unit sphere~$S^{2}$~\cite{Hatcher2001,dryden2005}.
Since the state mapping to impossible always contained in a plane,
$\bar{f}^{-1}\left(\imposs\right)$ in the above case 2 - 4 does not
contain any area in $S^{2}$. In contrast, if $\bar{f}^{-1}\left(\imposs\right)$
contains some area in $S^{2}$, it must be the situation of case 1.
Furthermore, the larger the area, the more 1-dimensional projectors
which is inconsistent with any particular Born rule probability measure.
Therefore, the area of $\bar{f}^{-1}\left(\imposs\right)$ could indicate
how much $\bar{f}$ deviate from the Born rule probability measures.\footnote{Notice that we only discuss only discuss real 1-dimensional projectors
in the Hilbert space of dimension 3. In general, we could discuss
all 1-dimensional projectors in any dimension. However, restricting
to $\mathbb{R}^{3}$ enable us to use cross product which will simplify
our discussion a lot, and $\mathbb{R}^{3}$ is easy to visualize and
imagine. Also, if a quantum IVPM~$\bar{\mu}$ has an empty core when
restricting to $\mathbb{R}^{3}$, $\bar{\mu}$ has an empty core on
the original space as well. So restricting on real 1-dimensional projectors
still provides valid results.}

In order to simplify the discussion, let's set the center of $\bar{f}^{-1}\left(\imposs\right)$
as the north pole, and assume $\bar{f}$ maps every state whose latitude
is larger than $\theta_{0}$ to impossible. In order to simplify the
notation, let's give the following definition.

>>>>

\noindent It is easy to establish that quantum interval-valued probability
measures generalize conventional quantum probability measures. In
particular, any quantum probability measure can be recast as an interval-valued
measure using the three intervals $\left[0,0\right]$, $\left[1,1\right]$
and \emph{$\left[0,1\right]$}, where $\left[0,0\right]$ and $\left[1,1\right]$
are called \imposs~and \necess~as before, and \emph{$\left[0,1\right]$}
is called \unknown~because it provides no information. Given a quantum
probability measure~$\mu:\events\rightarrow\left[0,1\right]$, we
define a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
by $\bar{\mu}(P)=\iota\left(\mu(P)\right)$, where $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is defined by 
\begin{equation}
\iota(x)=\begin{cases}
\necess & \textrm{if }x=1\textrm{ ;}\\
\imposs & \textrm{if }x=0\textrm{ ;}\\
\unknown & \textrm{otherwise.}
\end{cases}\label{eq:homomorphism}
\end{equation}
This measure represents the beliefs of an experimenter with no prior
knowledge about the particular quantum system in question. Formally,
we can ask: what can we deduce about the state of a quantum system
given a quantum interval-valued probability measure, i.e., given observations
done with finite resources. In the case the intervals are infinitely
precise the question reduces to Gleason's theorem which states that
the state of the quantum system is uniquely determined by the probability
measure. But surely the less resources are available, the less precise
the intervals, and the less we expect to know about the state of the
system. To formally state and answer this question we begin with defining
the \emph{core} and \emph{convexity} of a probability measure as follows.

@@

To prove the other direction, consider the linear transformation~$T:\mathbb{C}^{3}\rightarrow\mathbb{C}^{2}$
defined by $T\left(\ket{\varphi}\right)=\begin{pmatrix}\ip{\psi}{\varphi}\\
\ip{\phi}{\varphi}
\end{pmatrix}$. Since $\ket{\psi}\nparallel\ket{\phi}$, we have $\mathrm{rank}\left(T\right)=2$.
Together with the rank equation~\cite{FraleighBeauregard1995}, we
have
\[
\dim\left(\ker T\right)=\mathrm{nullity}\left(T\right)=3-\mathrm{rank}\left(T\right)=3-2=1\textrm{ .}
\]
By the other direction of the proof, we have already known $\ket{\psi}\times\ket{\phi}\in\ker T$.
Therefore, $\ket{\varphi}\parallel\left(\ket{\psi}\times\ket{\phi}\right)$.
\end{proof}
Since interval-valued probability measures~$\bar{\mu}$ and interval-valued
frame functions~$\bar{f}$ are one-to-one correspondence, studying
1-dimensional projectors mapping to impossible by $\bar{\mu}$ is
the same as studying the points on $\mathbb{C}S^{2}$ mapping to impossible
by $\bar{f}$. These points, denoted by $\bar{f}^{-1}\left(\imposs\right)=\set{\ket{\psi}\in\mathbb{C}^{3}}{\bar{f}(\ket{\psi})=\imposs}$,
play an important role to understand $\bar{f}$. There are 4 different
cases when analyzing $\bar{f}^{-1}\left(\imposs\right)$:
\begin{enumerate}
\item $\bar{f}^{-1}\left(\imposs\right)=\emptyset$. If $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, we could say nothing about $\ket{\phi}$.
\item $\bar{f}^{-1}\left(\imposs\right)=\left\{ \ket{\psi_{0}}\right\} $.
If $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$ for some pure
state~$\ket{\phi}$, $\ket{\phi}$ could be any state orthogonal
to $\ket{\psi_{0}}$.
\item Consider $\left\{ \ket{\psi_{0}},\ket{\psi_{1}}\right\} \subseteq\bar{f}^{-1}\left(\imposs\right)$.
If $\left(\ket{\psi_{0}}\times\ket{\psi_{1}}\right)\parallel\left(\ket{\psi_{0}'}\times\ket{\psi_{1}'}\right)$
for any $\left\{ \ket{\psi_{0}'},\ket{\psi_{1}'}\right\} \subseteq\bar{f}^{-1}\left(\imposs\right)$,
then $\ket{\psi_{0}}\times\ket{\psi_{1}}$ is the only pure state
which might induce $\bar{f}$. In another word, if $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, then $\ket{\phi}\parallel\left(\ket{\psi_{0}}\times\ket{\psi_{1}}\right)$.
For example, example~\ref{ex:three-dimensional-4-value} is in this
case when the core is empty; an interval-valued frame function defined
by $\bar{f}\left(\ket{\psi}\right)=\iota\left(f_{\phi}^{B}\left(\ket{\psi}\right)\right)$
is also in this case when the core is the singleton set~$\left\{ f_{\phi}^{B}\right\} $,
where $\iota$ is defined in equation~(\ref{eq:homomorphism}).
\item If there exist two pair of states and $\left\{ \ket{\psi_{0}},\ket{\psi_{1}}\right\} $
and $\left\{ \ket{\psi_{0}'},\ket{\psi_{1}'}\right\} \subseteq\bar{f}^{-1}\left(\imposs\right)$
such that $\left(\ket{\psi_{0}}\times\ket{\psi_{1}}\right)\nparallel\left(\ket{\psi_{0}'}\times\ket{\psi_{1}'}\right)$,
then both $\ket{\psi_{0}}\times\ket{\psi_{1}}$ and $\ket{\psi_{0}'}\times\ket{\psi_{1}'}$
might induce $\bar{f}$. However, neither $f_{\psi_{0}\times\psi_{1}}^{B}$
nor $f_{\psi_{0}'\times\psi_{1}'}^{B}$ is in the core of $\bar{f}$
because $f_{\psi_{0}\times\psi_{1}}^{B}$ is inconsistent with $\bar{f}\left(\ket{\psi_{0}'}\right)=\bar{f}\left(\ket{\psi_{1}'}\right)=\imposs$
and $f_{\psi_{0}'\times\psi_{1}'}^{B}$ is inconsistent with $\bar{f}\left(\ket{\psi_{0}}\right)=\bar{f}\left(\ket{\psi_{1}}\right)=\imposs$.
Actually, $\bar{f}$ has an empty core, and this is the case for example~\ref{ex:three-dimensional-three-value}.
\end{enumerate}
In summary, if we have an idea of $\bar{f}^{-1}\left(\imposs\right)$,
we could have some idea of the pure states in core.

If $\bar{f}$ has a non-empty core, $\bar{f}$ must be in the above
case 1 - 3 so that $\bar{f}^{-1}\left(\imposs\right)\cap S^{2}$ contains
no area in $S^{2}$.\footnote{Notice that we only discuss the part of $\bar{f}^{-1}\left(\imposs\right)$
which is in $S^{2}$. In general, we could discuss the whole $\bar{f}^{-1}\left(\imposs\right)\subseteq\mathbb{C}S^{2}$.
However, restricting to $S^{2}$ enable us to visualize and imagine
easily. Also, if a interval-valued frame function~$\bar{f}$ has
an empty core when restricting to $S^{2}$, $\bar{\mu}$ has an empty
core on the original space as well. So restricting to $S^{2}$ still
provides valid results.} In contrast, if $\bar{f}^{-1}\left(\imposs\right)\cap S^{2}$ contains
some area in $S^{2}$, it must be the situation of case 4, and $\bar{f}$
has an empty core. Furthermore, the larger the area, the more state
vectors which are inconsistent with any particular Born rule frame
function. Therefore, the area of $\bar{f}^{-1}\left(\imposs\right)\cap S^{2}$
could indicate how much $\bar{f}$ deviate from the Born rule frame
functions.

In order to simplify the discussion, consider the center of $\bar{f}^{-1}\left(\imposs\right)\cap S^{2}$
as the north pole, and assume $\bar{f}$ maps every state whose latitude
is larger than $\theta_{0}$ to impossible.
